---
title: "Distinct locals and visitors"
date: "last compiled date: `r format(Sys.time(), '%d %B, %Y')`"
output:
  workflowr::wflow_html:
    toc: false
    code_folding: show
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(tidyverse)
library(sf)
library(here)
library(purrrogress)
```

Locals are users with homes within that specific location or hexagonal cell. Conversely, visitors are all other users with homes outside of the cell.

## Load data 

```{r}
# load aggregated grids 
grids <- read_sf(here("data/derived_data/spatial_hex_grid.shp"), quiet = T) %>% 
  st_transform(crs = 3414)

# load twitter dataset 
df <- readRDS(here("data/derived_data/deidentified_sg_tweets_updated.rds")) 

# load identified home locations 
identified_hms <- read_csv(here("data/derived_data/identified_hms.csv"))
```


## Distinct locals and visitors 

We only consider users with identified home locations. For each grid, a user is assigned as 'local' or 'visitor' according to his/her home location. 

- **Local**: If the visited grid is the same as the home location of the user, the user is assigned as 'local'. 
- **Visitor**: If the visited grid is different from the home location of the user, the user is assigned as 'visitor'. 

The distinct users in grids are saved under `data/derived_data/` directory. 

```{r}
# users in each of grids 
users_in_grids <- df %>% dplyr::select(-created_at) %>% distinct(u_id, grid_id)

# distinct users 
distinct_users_in_grid <- function(df_users_in_grids, df_home, grid_index){
  output <- df_users_in_grids %>% 
    filter(grid_id == grid_index) %>% 
    left_join(., df_home) %>% 
    drop_na() %>% #remove users without identified home location
    mutate(type = if_else(grid_id == home, "local", "visitor"))
  return(output)
}

if(file.exists(here("data/derived_data/distinct_users.rds"))){
  distinct_users <- readRDS(here("data/derived_data/distinct_users.rds"))
}else{
  distinct_users <- do.call(bind_rows, map(grids$grid_id, with_progress(function(x) distinct_users_in_grid(users_in_grids, identified_hms, x))))
  saveRDS(distinct_users, file = here("data/derived_data/distinct_users.rds"))
}

head(distinct_users)
```






