---
title: "Flow Mapping"
author: "Qingqing Chen"
date: "Last compiled date: `r format(Sys.time(), '%d %B, %Y')`"
output: github_document
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align='center', fig.width = 10, fig.height = 10, warning = FALSE, message = FALSE)
library(tidyverse)
library(sf)
library(tmap)
library(purrrogress)
library(ggspatial)
library(here)
source(here("R/flow_mapping.R"))
```

<style>
body {text-align: justify}
</style>

Flows into or out of a neighbourhood can happen in all cardinal directions and with different intensities and distances. Visualizing the entire network of these connections is not very insightful for making neighbourhood-level inferences. This is why we create a single metric that summarizes the overall direction and strength of a flow to or from a neighbourhood, inspired by Tobler's (1981) work on vector fields to display flow patterns.

We first aggregate the sectors of a grid into four zones, with each zone corresponding to a specific angle and vector, relative to ‘East’ (0°): the North East (NE) zone (45°), North West (NW) zone (135°), South West (SW) zone (225°), and South East (SE) zone (315°). This simplification alleviates computational complexity while maintaining the overall flow orientation of a neighbourhood. For each vector, its direction represents where visitors come from or where locals go out to, and its magnitude indicates the number of people in that flow. The resultant vector - the sum of the four vectors - is used to represent the overall flow direction of each grid. To visualize all resulting vectors uniformly on a single map, we log-transform the length of each vector (to prevent some arrows from being much longer than others). Furthermore, we also represent the number of users active in the grid with the width of the resultant vector arrows.

## Load data
```{r}
# sg boundary
sg_boundary <- readRDS(here("data/derived_data/sg_boundary.rds"))
# hexagonal grids 
grids <- read_sf(here("data/derived_data/spatial_hex_grid.shp"), quiet = T) %>%
  st_transform(crs = 3414)
# grid centroids 
grid_centroids <- grids %>% st_centroid()

# divided sectors 
grid_sectors <- readRDS(here("data/derived_data/grid_sectors.rds"))

# inflow and outflow distance
dist_visitors <- readRDS(here("data/derived_data/dist_visitor_points.rds"))
dist_locals <- readRDS(here("data/derived_data/dist_local_points.rds"))

# inflow and outflow diversity
inflow_diversity <- readRDS(here("data/derived_data/inflow_diversity.rds"))
outflow_diversity <- readRDS(here("data/derived_data/outflow_diversity.rds")) 
```

## Vector flow analysis

### Inflow vectors

```{r}
if(file.exists(here("data/derived_data/inflow_vectors.rds"))){
  inflow_vectors <- readRDS(here("data/derived_data/inflow_vectors.rds"))
}else{
  dist_visitors_nest <- dist_visitors %>% 
  nest(data = c(u_id, home, created_at, dist_hm2grid, dist_hm2grid_km))
  inflow_vectors <- do.call(bind_rows, map2(dist_visitors_nest$grid_id, dist_visitors_nest$data, with_progress(cal_inflow_vectors)))
  saveRDS(inflow_vectors, file = here("data/derived_data/inflow_vectors.rds"))
}

# scale inflow vectors
inflow_vectors_scaled <- inflow_vectors %>% 
  mutate(magnitude_rescaled = scales::rescale(log(magnitude))) %>% 
  mutate(start_x = X + 1 + magnitude_rescaled * 1000 * cos(angle),
         start_y = Y + 1 + magnitude_rescaled * 1000 * sin(angle),
         total_user = scales::rescale(log(total_user)))
head(inflow_vectors_scaled)
```


### Outflow vectors 

```{r}
if(file.exists(here("data/derived_data/outflow_vectors.rds"))){
  outflow_vectors <- readRDS(here("data/derived_data/outflow_vectors.rds"))
}else{
  dist_locals_nest <- dist_locals %>% 
  nest(data = c(u_id, grid_id, created_at, dist_hm2grid, dist_hm2grid_km)) 
  outflow_vectors <- do.call(bind_rows, map2(dist_locals_nest$home, dist_locals_nest$data, with_progress(cal_outflow_vectors)))
  saveRDS(outflow_vectors, file = here("data/derived_data/outflow_vectors.rds"))
}

# scale outflow vectors 
outflow_vectors_scaled <- outflow_vectors %>%
  mutate(magnitude_rescaled = scales::rescale(log(magnitude)),
         end_x = X + 1 + magnitude_rescaled * 1000 * cos(angle),
         end_y = Y + 1 + magnitude_rescaled * 1000 * sin(angle),
         total_user = scales::rescale(log(total_user)))
head(outflow_vectors_scaled)
```


## Flow map visualization 

### Background preparation 

#### Background: Travel distance

```{r}
# mean inflow distance 
bg_dist_inflow <- dist_visitors %>% 
  group_by(grid_id, u_id) %>% 
  dplyr::summarise(mean_dist_user = mean(dist_hm2grid_km)) %>% 
  group_by(grid_id) %>% 
  dplyr::summarise(mean_dist_grid = mean(mean_dist_user),
                   sd_dist_grid = sd(mean_dist_user)) %>% 
  mutate(cut_mean_dist_grid = cut(mean_dist_grid, 
                                  breaks=c(quantile(mean_dist_grid, probs = seq(0, 1, by = 0.20))), include.lowest=TRUE), 
         cut_sd_dist_grid = cut(sd_dist_grid, 
                                breaks = c(quantile(sd_dist_grid, probs = seq(0, 1, by = 0.2))), include.lowest = TRUE)) %>% 
  left_join(., grids) %>% 
  st_as_sf()

# mean outflow distance 
bg_dist_outflow <- dist_locals %>% 
  group_by(home, u_id) %>% 
  dplyr::summarise(mean_dist_user = mean(dist_hm2grid_km)) %>% 
  group_by(home) %>% 
  dplyr::summarise(mean_dist_grid = mean(mean_dist_user), 
                   sd_dist_grid = sd(mean_dist_user)) %>% 
  mutate(cut_mean_dist_grid = cut(mean_dist_grid, 
                                  breaks=c(quantile(mean_dist_grid, probs = seq(0, 1, by = 0.20))), include.lowest=TRUE), 
         cut_sd_dist_grid = cut(sd_dist_grid, 
                                breaks = c(quantile(sd_dist_grid, probs = seq(0, 1, by = 0.2))), include.lowest = TRUE)) %>% 
  left_join(., grids, by = c("home" = "grid_id")) %>% 
  st_as_sf()
```


#### Background: Neighbourhood ratio
```{r}
if(file.exists(here("data/derived_data/neigh_ratio_inflow.rds"))){
  bg_neighRatio_inflow <- readRDS(here("data/derived_data/neigh_ratio_inflow.rds")) %>% 
    dplyr::select(grid_id, type, ratio_divide_sumWeigh) %>% 
    na.omit() %>% 
    spread(key = "type", value = "ratio_divide_sumWeigh") %>% 
    replace(., is.na(.), 0) %>% 
    left_join(., grids) %>% 
    st_as_sf(crs = 3414) %>% 
    mutate(ratio = cut(inner, breaks=c(quantile(inner, probs = seq(0, 1, by = 0.20))), include.lowest=TRUE))
}else{
  dist_visitors_weight <- dist_visitors %>% 
  dplyr::select(grid_id, u_id, home, created_at) %>% 
  group_by(u_id) %>% 
  mutate(total_tweets = n(), weight = 1/total_tweets) %>% 
  ungroup()
  
  neigh_ratio_inflow <- do.call(rbind, map(dist_visitors_nest$grid_id, with_progress(function(x) cal_neigh_ratio(dist_visitors_weight, x, user_type = "visitor"))))
  saveRDS(neigh_ratio_inflow, file = here("data/derived_data/neigh_ratio_inflow.rds"))
}

if(file.exists(here("data/derived_data/neigh_ratio_outflow.rds"))){
  bg_neighRatio_outflow <- readRDS(here("data/derived_data/neigh_ratio_outflow.rds")) %>% 
    dplyr::select(home, type, ratio_divide_sumWeigh) %>% 
    na.omit() %>% 
    spread(key = "type", value = "ratio_divide_sumWeigh") %>% 
    replace(., is.na(.), 0) %>% 
    left_join(., grids, by = c("home" = "grid_id")) %>% 
    st_as_sf(crs = 3414) %>% 
    mutate(ratio = cut(inner, breaks=c(quantile(inner, probs = seq(0, 1, by = 0.20))), include.lowest=TRUE))
}else{
  dist_locals_weight <- dist_locals %>% 
  dplyr::select(home, u_id, grid_id, created_at) %>% 
  group_by(u_id) %>% 
  mutate(total_tweets = n(), weight = 1/total_tweets) %>% 
  ungroup()
  
  neigh_ratio_outflow <- do.call(rbind, map(dist_locals_nest$home, with_progress(function(x) cal_neigh_ratio(dist_locals_weight, x, user_type = "local"))))
  saveRDS(neigh_ratio_outflow, file = here("data/derived_data/neigh_ratio_outflow.rds"))
}
```

#### Background: Diversity

```{r}
bg_div_inflow <-  inflow_diversity %>% 
  dplyr::select(grid_id, norm_div_shannon) %>% 
   mutate(cut_div = cut(norm_div_shannon, breaks=c(quantile(norm_div_shannon, probs = seq(0, 1, by = 0.20))), include.lowest=TRUE))

bg_div_outflow <-  outflow_diversity %>% 
  dplyr::select(grid_id, norm_div_shannon) %>% 
  mutate(cut_div = cut(norm_div_shannon, breaks=c(quantile(norm_div_shannon, probs = seq(0, 1, by = 0.20))), include.lowest=TRUE))
```



### Flow maps 

#### Inflow with neighbourhood ratio background 

```{r}
viz_flows(bg_neighRatio_inflow, inflow_vectors_scaled,
          bg.fill.var = "ratio", 
          quantile.var = "inner",
          palette = "OrRd",
          start.x = "start_x", start.y = "start_y", 
          end.x = "X", end.y = "Y", 
          title = expression(paste("(a) ", bold("Direction of Flows: "), "background is neighbourhood ratio")), 
          legend.nm = "Neighborhood ratio")
```

#### Outflow with neighbourhood ratio background 
```{r}
viz_flows(bg_neighRatio_outflow, outflow_vectors_scaled, 
          quantile.var = "inner",
          palette = "Purples",
          bg.fill.var = "ratio", 
          start.x = "X", start.y = "Y", 
          end.x = "end_x", end.y = "end_y",
          title = expression(paste("(a) ", bold("Direction of Flows: "), "background is neighbourhood ratio")), 
          legend.nm = "Neighborhood ratio")
```


#### Inflow with average distance background 

```{r}
viz_flows(bg_dist_inflow, inflow_vectors_scaled,
          bg.fill.var = "cut_mean_dist_grid",
          quantile.var = "mean_dist_grid",
          palette = "OrRd",
          start.x = "start_x", start.y = "start_y",
          end.x = "X", end.y = "Y",
          title = expression(paste("(b) ", bold("Direction of Flows: "), "background is average incoming distance")),
          legend.nm  = "Avg.distance (km)")
```

### Outflow average distance background 
```{r}
viz_flows(bg_dist_outflow, outflow_vectors_scaled, 
          bg.fill.var = "cut_mean_dist_grid", 
          quantile.var = "mean_dist_grid", 
          palette = "Purples",
          start.x = "X", start.y = "Y", 
          end.x = "end_x", end.y = "end_y", 
          title = expression(paste("(b) ", bold("Direction of Flows: "), "background is average outgoing distance")),
          legend.nm = "Avg.distance (km)")
```


#### Inflow with S.D.distance background
```{r}
viz_flows(bg_dist_inflow, inflow_vectors_scaled, 
          bg.fill.var = "cut_sd_dist_grid", 
          quantile.var = "sd_dist_grid",
          palette = "OrRd",
          start.x = "start_x", start.y = "start_y", 
          end.x = "X", end.y = "Y", 
          title = expression(paste("(c) ", bold("Direction of Flows: "), "background is S.D. of incoming distance")),
          legend.nm = "S.D.distance(km)")
```


#### Outflow with S.D.distance background
```{r}
viz_flows(bg_dist_outflow, outflow_vectors_scaled, 
          bg.fill.var = "cut_sd_dist_grid", 
          quantile.var = "sd_dist_grid",
          palette = "Purples",
          start.x = "X", start.y = "Y", 
          end.x = "end_x", end.y = "end_y", 
          title = expression(paste("(c) ", bold("Direction of Flows: "), "background is S.D. of outgoing distance")),
          legend.nm = "S.D.distance(km)")
```

#### Inflow with diversity background 

```{r}
viz_flows(bg_div_inflow %>% filter(grid_id %in% inflow_vectors_scaled$grid_id), inflow_vectors_scaled,
          bg.fill.var = "cut_div", 
          quantile.var = "norm_div_shannon",
          palette = "OrRd",
          start.x = "start_x", start.y = "start_y", 
          end.x = "X", end.y = "Y", 
          title = expression(paste("(d) ", bold("Direction of Flows: "), "background is normalized inflow diversity")),
          legend.nm = "Norm.diversity")
```

#### Outflow with diversity background 

```{r}
viz_flows(bg_div_outflow %>% filter(grid_id %in% outflow_vectors_scaled$home_id),  outflow_vectors_scaled,
          bg.fill.var = "cut_div", 
          quantile.var = "norm_div_shannon",
           palette = "Purples",
          start.x = "X", start.y = "Y", 
          end.x = "end_x", end.y = "end_y", 
          title = expression(paste("(d) ", bold("Direction of Flows: "), "background is normalized outflow diversity")),
          legend.nm = "Norm.diversity")
```
