---
title: "Regression analysis"
author: "Qingqing Chen"
date: "Last compiled date: `r format(Sys.time(), '%d %B, %Y')`"
output: github_document
editor_options:
  chunk_output_type: console
---

<style>
body {text-align: justify}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align='center', fig.width = 10, fig.height = 10, warning = FALSE, message = FALSE)
library(tidyverse)
library(sf)
library(lubridate)
library(tmap)
library(stars)
library(RColorBrewer)
library(purrrogress)
library(here)
library(broom)
library(spdep)
library(PerformanceAnalytics)
library(stargazer)
source(here("R/viz.R"))
```

To contextualize the mobility patterns in our analysis, we compare two of the aforementioned metrics – distance and diversity – with the socio-economic and built environment characteristics of each neighbourhood through a spatial regression analysis. We do so by drawing on a set of publicly available datasets, including the HDB Resale Flat Prices data, HDB Property Information data, SLA Street Directory data, and LTA Train Station data. These datasets allow us to create six independent variables shown as follows:

  - Mean resale price per square meter;
  - Percentage of 1-Room and 2-Room rental flats;
  - Percentage of residential building area;
  - Percentage of industrial building area;
  - Distance to the nearest MRT station;
  - Distance to the central grid cell.
  
We aggregate the variables to individual neighbourhoods (i.e. with the same spatial resolution as neighbourhoods) and only keep neighbourhoods that have public housing. It is important to stress that we employ this regression model not to predict travel distance or mobility diversity based on these independent variables but rather to provide an initial quantitative exploration of the potential underlying covariates that influence mobility patterns in Singapore.

## Load data 

```{r}
# hexagonal grids 
grids <- read_sf(here("data/derived_data/spatial_hex_grid.shp"), quiet = T) %>%
  st_transform(crs = 3414)

# grid centroids 
grid_centroids <- grids %>% st_centroid()

# HDB buildings
hdb_sf <-  read_sf(here("data/derived_data/spatial_hdb_building.shp"))

# inflow and outflow distance
dist_inflow <- readRDS(here("data/derived_data/dist_visitor_points.rds"))
dist_outflow <- readRDS(here("data/derived_data/dist_local_points.rds"))

# inflow and outflow diversity 
div_inflow <- readRDS(here("data/derived_data/inflow_diversity.rds"))
div_outflow <- readRDS(here("data/derived_data/outflow_diversity.rds"))
```


## Dependent variables 

### HDB resale price
```{r}
if(file.exists(here("data/derived_data/sf_resale.rds"))){
  sf_resale <- readRDS(here("data/derived_data/sf_resale.rds"))
}else{
  df_resale_12to14 <- read_csv(here("data/raw_data/resale-flat-prices/resale-flat-prices-based-on-registration-date-from-mar-2012-to-dec-2014.csv"))
  df_resale_15to16  <- read_csv(here("data/raw_data/resale-flat-prices/resale-flat-prices-based-on-registration-date-from-jan-2015-to-dec-2016.csv"))  
  df_resale <- bind_rows(df_resale_12to14, df_resale_15to16) %>% 
    mutate(month = ymd(month, truncated = 1)) %>% 
    filter(month >= '2012-07-01' & month < '2016-11-01') 
  # format street name 
  df_resale <- df_resale %>% 
    mutate(ROAD_NAME = gsub("\\bAVE\\b", "AVENUE", street_name),
           ROAD_NAME = gsub("\\bBT\\b", "BUKIT", ROAD_NAME),
           ROAD_NAME = gsub("\\bCL\\b", "CLOSE", ROAD_NAME),
           ROAD_NAME = gsub("\\bCRES\\b", "CRESCENT", ROAD_NAME),
           ROAD_NAME = gsub("\\bCTRL\\b", "CENTRAL", ROAD_NAME),
           ROAD_NAME = gsub("\\bC'WEALTH\\b", "COMMONWEALTH", ROAD_NAME),
           ROAD_NAME = gsub("\\bCTR\\b", "CENTRE", ROAD_NAME),
           ROAD_NAME = gsub("\\bDR\\b", "DRIVE", ROAD_NAME),
           ROAD_NAME = gsub("\\bGDNS\\b", "GARDENS", ROAD_NAME),
           ROAD_NAME = gsub("\\bHTS\\b", "HEIGHTS", ROAD_NAME),
           ROAD_NAME = gsub("\\bJLN\\b", "JALAN", ROAD_NAME),
           ROAD_NAME = gsub("\\bKG\\b", "KAMPONG", ROAD_NAME),
           ROAD_NAME = gsub("\\bLOR\\b", "LORONG", ROAD_NAME),
           ROAD_NAME = gsub("\\bMKT\\b", "MARKET", ROAD_NAME),
           ROAD_NAME = gsub("\\bNTH\\b", "NORTH", ROAD_NAME),
           ROAD_NAME = gsub("\\bPL\\b", "PLACE", ROAD_NAME),
           ROAD_NAME = gsub("\\bPK\\b", "PARK", ROAD_NAME),
           ROAD_NAME = gsub("\\bRD\\b", "ROAD", ROAD_NAME),
           ROAD_NAME = gsub("\\bST\\b", "STREET", ROAD_NAME),
           ROAD_NAME = gsub("\\bSTH\\b", "SOUTH", ROAD_NAME),
           ROAD_NAME = gsub("\\bTER\\b", "TERRACE", ROAD_NAME),
           ROAD_NAME = gsub("\\bTG\\b", "TANJONG", ROAD_NAME),
           ROAD_NAME = gsub("\\bUPP\\b", "UPPER", ROAD_NAME)) %>%
      mutate(ROAD_NAME = case_when(
        ROAD_NAME == "STREET. GEORGE'S LANE" ~ "SAINT GEORGE'S LANE",
        ROAD_NAME == "STREET. GEORGE'S ROAD" ~ "SAINT GEORGE'S ROAD",
        TRUE ~ ROAD_NAME
      )) %>%
      dplyr::rename(HOUSE_BLK_ = block) %>%
      unite(ROAD_BLK, c("ROAD_NAME", "HOUSE_BLK_")) %>%
      mutate(ROAD_BLK = case_when(
        ROAD_BLK == "CHOA CHU KANG AVENUE 2_297A" ~ "CHOA CHU KANG AVENUE 2_297",
        ROAD_BLK == "CHOA CHU KANG AVENUE 2_297B" ~ "CHOA CHU KANG AVENUE 2_297",
        TRUE ~ ROAD_BLK
      ))
    
   # add geometry to each resale transaction
    sf_resale <- hdb_sf %>%
      left_join(df_resale, ., by = c("ROAD_BLK" = "ROAD_BLK")) %>%
      st_as_sf() %>%
      st_transform(crs = 3414) %>%
      st_join(., grids, largest = T) %>% 
      filter(!is.na(grid_id)) 
    saveRDS(sf_resale, file = here("data/derived_data/sf_resale.rds"))
}

head(sf_resale)
```

```{r}
mean_resale <- sf_resale %>% 
  st_set_geometry(NULL) %>% 
  dplyr::select(grid_id, floor_area_sqm, resale_price) %>% 
  mutate(resale_price_sqm = resale_price/floor_area_sqm) %>% 
  group_by(grid_id) %>% 
  dplyr::summarise(mean_resale_price_sqm = mean(resale_price_sqm),
                   sd_resale_price_sqm = sd(resale_price_sqm)) 

## grid cells with available socio-economic variables and human mobility indicators
qualified_grids_inflow <- intersect(mean_resale$grid_id, div_inflow$grid_id) %>% intersect(dist_inflow$grid_id)
qualified_grids_outflow <- intersect(mean_resale$grid_id, div_outflow$grid_id) %>% intersect(dist_outflow$grid_id)

mean_resale_sf <- mean_resale %>%
  filter(grid_id %in% qualified_grids_inflow) %>% 
  left_join(., grids, by = c("grid_id" = "grid_id")) %>% 
  st_sf()
```


```{r}
spatial_viz(mean_resale_sf, 
            fill_var = "mean_resale_price_sqm", 
            legend_title = "Avg.resale price/sqm", 
            main_title = "(a) Spatial distribution of average resale price", 
            main.title_size = 1.4,
            legend.hist_height = 0.25, legend.hist_width = 0.5,
            legend_width = 0.4, legend.hist_size = 0.5,
            legend.title_size = 1.2, legend.text_size = 0.65,
            digits = 0, palette = "PuRd")

# spatial_viz(mean_resale_sf, 
#             fill_var = "sd_resale_price_sqm", 
#             legend_title = "S.D.resale price/sqm", 
#             main_title = "(a) Spatial distribution of S.D of resale price", 
#             main.title_size = 1.4,
#             legend.hist_height = 0.25, legend.hist_width = 0.5,
#             legend_width = 0.4, legend.hist_size = 0.5,
#             legend.title_size = 1.2, legend.text_size = 0.65,
#             digits = 0, palette = "PuRd")
```

```{r fig.height=5}
violin_viz(mean_resale_sf, var = "mean_resale_price_sqm", 
           labs.x = "", labs.y = "Avg.resale price/sqm", breaks = seq(0, 9000, 1000),
           y.shift = 1050, x.shift = 1.4, text.size = 4, digits = 0)
```

### Percentage of 1-ROOM and 2-Room rental flats

```{r}
#HDB Property Information: https://data.gov.sg/dataset/hdb-property-information
if(file.exists(here("data/derived_data/df_pct_1_2room_rental.rds"))){
  df_pct_1_2room_rental <- readRDS(here("data/derived_data/df_pct_1_2room_rental.rds"))
}else{
  df_hdb_property <- read_csv(here("data/raw_data/hdb-property-information.csv")) 
  df_hdb_property_updated <-  df_hdb_property %>% 
    mutate(ROAD_NAME = gsub("\\bDR\\b", "DRIVE", street),
           ROAD_NAME = gsub("\\bRD\\b", "ROAD", ROAD_NAME),
           ROAD_NAME = gsub("\\bCRES\\b", "CRESCENT", ROAD_NAME),
           ROAD_NAME = gsub("\\bAVE\\b", "AVENUE", ROAD_NAME),
           ROAD_NAME = gsub("\\bST\\b", "STREET", ROAD_NAME),
           ROAD_NAME = gsub("\\bCTRL\\b", "CENTRAL", ROAD_NAME),
           ROAD_NAME = gsub("\\bNTH\\b", "NORTH", ROAD_NAME),
           ROAD_NAME = gsub("\\bSTH\\b", "SOUTH", ROAD_NAME),
           ROAD_NAME = gsub("\\bBT\\b", "BUKIT", ROAD_NAME),
           ROAD_NAME = gsub("\\bC'WEALTH\\b", "COMMONWEALTH", ROAD_NAME),
           ROAD_NAME = gsub("\\bCL\\b", "CLOSE", ROAD_NAME),
           ROAD_NAME = gsub("\\bPK\\b", "PARK", ROAD_NAME),
           ROAD_NAME = gsub("\\bJLN\\b", "JALAN", ROAD_NAME),
           ROAD_NAME = gsub("\\bKG\\b", "KAMPONG", ROAD_NAME),
           ROAD_NAME = gsub("\\bPL\\b", "PLACE", ROAD_NAME),
           ROAD_NAME = gsub("\\bLOR\\b", "LORONG", ROAD_NAME),
           ROAD_NAME = gsub("\\bTER\\b", "TERRACE", ROAD_NAME),
           ROAD_NAME = gsub("\\bMKT\\b", "MARKET", ROAD_NAME),
           ROAD_NAME = gsub("\\bUPP\\b", "UPPER", ROAD_NAME),
           ROAD_NAME = gsub("\\bGDNS\\b", "GARDENS", ROAD_NAME),
           ROAD_NAME = gsub("\\bTG\\b", "TANJONG", ROAD_NAME),
           ROAD_NAME = gsub("\\bHTS\\b", "HEIGHTS", ROAD_NAME),
           ROAD_NAME = gsub("\\bCTR\\b", "CENTRE", ROAD_NAME)) %>%
    mutate(ROAD_NAME = case_when(
          ROAD_NAME == "STREET. GEORGE'S LANE" ~ "SAINT GEORGE'S LANE",
          ROAD_NAME == "STREET. GEORGE'S ROAD" ~ "SAINT GEORGE'S ROAD",
          ROAD_NAME == "SECTOR A SIN MING IND EST" ~ "SECTOR A SIN MING INDUSTRIAL ESTATE", 
          TRUE ~ ROAD_NAME
          ))  %>%
    dplyr::rename(HOUSE_BLK_ = blk_no) %>%
    unite(ROAD_BLK, c("ROAD_NAME", "HOUSE_BLK_")) %>% 
    dplyr::select(ROAD_BLK, `1room_rental`, `2room_rental`, total_dwelling_units) %>% 
    mutate(ROAD_BLK = case_when(
            ROAD_BLK == "CHOA CHU KANG AVENUE 2_297A" ~ "CHOA CHU KANG AVENUE 2_297",
            ROAD_BLK == "CHOA CHU KANG AVENUE 2_297B" ~ "CHOA CHU KANG AVENUE 2_297",
            TRUE ~ ROAD_BLK
          ))
      
  # get extra two Blks that are under 'S' and 'K' building type
  extra_building_sf <- df_building %>% 
    filter(ROAD_NAME %in% c("NEW MARKET ROAD", "JALAN KUKOH") & HOUSE_BLK_ %in% c(1, 32)) %>% 
    unite(ROAD_BLK, c("ROAD_NAME", "HOUSE_BLK_")) %>% 
    st_transform(crs = 3414) %>% 
    dplyr::select(names(hdb_sf))

  df_1_2room_rental <- df_hdb_property_updated %>% 
    filter(total_dwelling_units != 0) %>% 
    left_join(., rbind(hdb_sf, extra_building_sf)) %>% 
    st_as_sf() %>% 
    st_make_valid() %>% 
    st_join(., grids, largest = T) %>% # join grids 
    filter(!is.na(grid_id)) %>% 
    group_by(grid_id) %>% 
    dplyr::summarise(`1room_rental` = sum(`1room_rental`), 
                     `2room_rental` = sum(`2room_rental`), 
                      total_dwelling_units = sum(total_dwelling_units))
    
  df_pct_1_2room_rental <- df_1_2room_rental %>% 
    st_set_geometry(NULL) %>% 
    mutate(pct_1_2room_rental = (`1room_rental` + `2room_rental`)/total_dwelling_units) 
  saveRDS(df_pct_1_2room_rental, file = here("data/derived_data/df_pct_1_2room_rental.rds"))
}
head(df_pct_1_2room_rental)
```


```{r}
df_pct_1_2room_rental_sf <- df_pct_1_2room_rental %>% 
              filter(grid_id %in% qualified_grids_inflow) %>% 
              left_join(., grids, by = c("grid_id" = "grid_id")) %>% 
              st_sf()

spatial_viz(df_pct_1_2room_rental_sf, 
            fill_var = "pct_1_2room_rental", 
            legend_title = "1&2 Room rentals (%)", 
            main_title = "(b) Spatial distribution of 1-Room and 2-Room rentals", 
            main.title_size = 1.4,
            legend.hist_height = 0.25, legend.hist_width = 0.5,
            legend_width = 0.4, legend.hist_size = 0.5,
            legend.title_size = 1.2, legend.text_size = 0.65, palette = "PuRd", 
            style = "fixed",
            breaks = c(0, 0.05, 0.1, 0.3, 0.5, 0.8))
```

```{r fig.height=5}
violin_viz(df_pct_1_2room_rental_sf,
           var = "pct_1_2room_rental",
           labs.x = "", labs.y = "1-Room & 2-Room Rentals (%)", 
           breaks = seq(0, 0.8, 0.1),
           y.shift = 0.13, x.shift = 1.4, text.size = 4)
```

### Landuse: Percentage of industrial building

```{r}
cal_building_area <- function(index, building_sf){
  # intersect with grids 
  intersect_grid <- building_sf %>% 
    filter(FEA_ID == index) %>% 
    st_make_valid() %>% 
    st_intersection(., grids) 
  # intersect area in grids
  intersect_grid %>% 
    st_set_geometry(NULL) %>% 
    dplyr::mutate(area_sqm = st_area(intersect_grid) %>% as.numeric()) %>% 
    dplyr::select(FEA_ID, grid_id, area_sqm)
}

if(file.exists(here("data/derived_data/industrial_areas_grids.rds"))){
  industrial_areas_grids <- readRDS(here("data/derived_data/industrial_areas_grids.rds"))
}else{
  #building type: I
  industrial_sf <- df_building %>% 
    filter(BLDG_TYPE_ == "I") %>% 
    filter(DATA_TYPE_ == "extg") %>% # Existing building outline
    st_transform(crs = 3414) %>% 
    unite(ROAD_BLK, c("ROAD_NAME", "HOUSE_BLK_"), sep = " ") 

  FEA_indexes <- industrial_sf$FEA_ID
  # industrial building areas in grids 
  industrial_areas_grids <- do.call(bind_rows, map(FEA_indexes, with_progress(function(x) cal_building_area(x, industrial_sf))))
  # percentage of commercial areas in grids 
  industrial_areas_grids <- industrial_areas_grids %>% 
    group_by(grid_id) %>% 
    dplyr::summarise(industrial_area_sqm = sum(area_sqm)) %>% 
    mutate(grid_area = st_area(grids[1, ]) %>% as.numeric(),
           pct_industrial_area = round(industrial_area_sqm/grid_area, 4)) 
  saveRDS(industrial_areas_grids, file = here("data/derived_data/industrial_areas_grids.rds"))
}

head(industrial_areas_grids)
```

```{r}
industrial_areas_grids_sf <- industrial_areas_grids %>% 
              filter(grid_id %in% qualified_grids_inflow) %>% 
              left_join(grids %>% filter(grid_id %in% qualified_grids_inflow), ., by = c("grid_id" = "grid_id")) %>% 
              replace(., is.na(.), 0) %>% 
              st_sf()

spatial_viz(industrial_areas_grids_sf, 
            fill_var = "pct_industrial_area", 
            legend_title = "Industrial area (%)", 
            main_title = "(c) Spatial distribution of industrial area", 
            main.title_size = 1.4,
            legend.hist_height = 0.25, legend.hist_width = 0.5,
            legend_width = 0.4, legend.hist_size = 0.5,
            legend.title_size = 1.2, legend.text_size = 0.65, palette = "PuRd", 
            style = "fixed",
            breaks = c(0, 0.05, 0.1, 0.15, 0.2, 0.25))
```


```{r fig.height=5}
violin_viz(industrial_areas_grids_sf,
           var = "pct_industrial_area",
           labs.x = "", labs.y = "Industrial area (%)", 
           breaks = seq(0, 0.25, 0.05), digits = 3,
           y.shift = 0.04, x.shift = 1.4, text.size = 4)
```

### Landuse: Percentage of residential building

```{r}
if(file.exists(here("data/derived_data/residential_areas_grids.rds"))){
  residential_areas_grids <- readRDS(here("data/derived_data/residential_areas_grids.rds"))
}else{
  #building type: C-Condominium, H-HDB Building, A-Apartment, E-Executive Condominium, R-Residential
  residential_sf <- df_building %>% 
    filter(BLDG_TYPE_ %in% c("C", "H", "A", "E", "R")) %>% 
    filter(DATA_TYPE_ == "extg") %>% # Existing building outline
    st_transform(crs = 3414) %>% 
    unite(ROAD_BLK, c("ROAD_NAME", "HOUSE_BLK_"), sep = " ") 

  FEA_indexes <- residential_sf$FEA_ID
  # residential areas in grids 
  residential_areas_grids <- do.call(bind_rows, map(FEA_indexes, with_progress(function(x) cal_building_area(x, residential_sf))))
  # percentage of residential areas in grids 
  residential_areas_grids <- residential_areas_grids %>% 
    group_by(grid_id) %>% 
    dplyr::summarise(residential_area_sqm = sum(area_sqm)) %>% 
    mutate(grid_area = st_area(grids[1, ]) %>% as.numeric(),
           pct_residential_area = round(residential_area_sqm/grid_area, 4)) 
  saveRDS(residential_areas_grids, file = here("data/derived_data/residential_areas_grids.rds"))
}

head(residential_areas_grids)
```


```{r}
residential_areas_grids_sf <- residential_areas_grids %>% 
              filter(grid_id %in% qualified_grids_inflow) %>% 
              left_join(., grids, by = c("grid_id" = "grid_id")) %>% 
  st_sf()

spatial_viz(residential_areas_grids_sf, 
            fill_var = "pct_residential_area", 
            legend_title = "Residential area (%)", 
            main_title = "(d) Spatial distribution of residential area", 
            main.title_size = 1.4,
            legend.hist_height = 0.25, legend.hist_width = 0.5,
            legend_width = 0.4, legend.hist_size = 0.5,
            legend.title_size = 1.2, legend.text_size = 0.65,
            palette = "PuRd")
```


```{r fig.height=5}
violin_viz(residential_areas_grids_sf, 
           var = "pct_residential_area", 
           labs.x = "", labs.y = "Residential area (%)", 
           breaks = seq(0, 0.32, 0.05),
           y.shift = 0.06, x.shift = 1.55, text.size = 4)
```

### Distance to the nearest MRT station 

```{r}
cal_dist2nearestMRT <- function(grid_index, grids, mrt){
  grid_centroid <- grids[grid_index, ] %>% st_centroid() 
  # create 20k buffer 
  grid_buffer <-  grid_centroid %>% st_buffer(20000) 
  # get MRT stations within the buffer 
  mrt_station_in_buffer <- grid_buffer %>% 
    st_join(mrt, .) %>%
    filter(!is.na(grid_id))
  
  # distance from grid centroid to MRT stations 
  dist2MRT <- st_distance(st_geometry(grid_centroid), st_geometry(mrt_station_in_buffer), by_element = TRUE) %>% as.numeric()
  
  # get the nearest distance 
  mrt_station_in_buffer %>% 
    st_set_geometry(NULL) %>% 
    mutate(dist2MRT_km = round(dist2MRT/1000, 4)) %>% 
    arrange(dist2MRT_km) %>% 
    slice(1)
}

if(file.exists(here("data/derived_data/dist2MRT_nearest.rds"))){
  dist2MRT_nearest <- readRDS(here("data/derived_data/dist2MRT_nearest.rds"))
}else{
  # MRT station
  mrt <- read_sf(here("data/raw_data/TrainStation_Jan2020/MRTLRTStnPtt.Shp"), quiet = T) %>% 
    st_transform(crs = 3414)
  dist2MRT_nearest <- do.call(bind_rows, map(grids$grid_id, with_progress(function(x) cal_dist2nearestMRT(x, grids, mrt))))
  saveRDS(dist2MRT_nearest, file = here("data/derived_data/dist2MRT_nearest.rds"))
}

head(dist2MRT_nearest)
```


```{r}
dist2MRT_nearest_sf <- dist2MRT_nearest %>% 
              filter(grid_id %in% qualified_grids_inflow) %>% 
              left_join(., grids, by = c("grid_id" = "grid_id")) %>% 
  st_sf()

spatial_viz(dist2MRT_nearest_sf, 
            fill_var = "dist2MRT_km", 
            legend_title = "Distance (km)", 
            main_title = "(e) Spatial distribution of distance to the nearest MRT station", 
            main.title_size = 1.4,
            legend.hist_height = 0.25, legend.hist_width = 0.5,
            legend_width = 0.4, legend.hist_size = 0.5,
            legend.title_size = 1.2, legend.text_size = 0.65,
            palette = "PuRd")
```


```{r fig.height=5}
violin_viz(dist2MRT_nearest_sf, 
           var = "dist2MRT_km", 
           labs.x = "", labs.y = "Distance (km)", 
           breaks = seq(0, 2.5, 0.5),
           y.shift = 0.5, x.shift = 1.4, text.size = 4)
```

### Distance from grid to 'center' grid 1232 - City Hall

```{r}
center_grid <- grid_centroids %>% filter(grid_id == 1232)
dist2_center_grid <- grid_centroids %>% 
  mutate(dist_grid2center = st_distance(., center_grid) %>% as.numeric(),
         dist_grid2center_km = dist_grid2center/1000) %>% 
  st_set_geometry(NULL)
head(dist2_center_grid)
```


```{r}
dist2_center_grid_sf <- dist2_center_grid %>% 
              filter(grid_id %in% qualified_grids_inflow) %>% 
              left_join(., grids, by = c("grid_id" = "grid_id")) %>% st_sf()

spatial_viz(dist2_center_grid_sf, 
            fill_var = "dist_grid2center_km", 
            legend_title = "Distance (km)", 
            main_title = "(f) Spatial distribution of distance to the central location", 
            main.title_size = 1.4,
            legend.hist_height = 0.25, legend.hist_width = 0.5,
            legend_width = 0.4, legend.hist_size = 0.5,
            legend.title_size = 1.2, legend.text_size = 0.65,
            palette = "PuRd")
```


```{r fig.height=5}
violin_viz(dist2_center_grid_sf, 
           var = "dist_grid2center_km", 
           labs.x = "", labs.y = "Distance (km)", 
           breaks = seq(0, 20, 5),
           y.shift = 5, x.shift = 1.55, text.size = 4)
```

## Dependent variables
### Travel distance

```{r}
# visitors travel distance 
mean_dist_inflow <- dist_inflow %>% 
  group_by(grid_id, u_id) %>% 
  dplyr::summarise(mean_dist_user = mean(dist_hm2grid_km)) %>% 
  group_by(grid_id) %>% 
  dplyr::summarise(mean_dist_km = mean(mean_dist_user)) %>% 
  mutate(user_type = "visitor") 

# locals travel distance 
mean_dist_outflow <- dist_outflow %>% 
  group_by(home, u_id) %>% 
  dplyr::summarise(mean_dist_user = mean(dist_hm2grid_km)) %>% 
  group_by(home) %>% 
  dplyr::summarise(mean_dist_km = mean(mean_dist_user)) %>% 
  dplyr::rename(grid_id = home) %>% mutate(user_type = "local")
```


```{r fig.height=8}
ggplot(bind_rows(mean_dist_inflow, mean_dist_outflow)) +
  geom_density(aes(mean_dist_km, fill = user_type, color = user_type), alpha = 0.5) +
  geom_vline(data = bind_rows(mean_dist_inflow, mean_dist_outflow) %>% group_by(user_type) %>% dplyr::summarise(mean = mean(mean_dist_km)), aes(xintercept = mean, color = user_type), size=1.5) +
  theme_bw() +
  theme(legend.position = c(0.9, 0.9)) +
  labs(x = "Avg.distance (km)", y = "Density", fill = "User type", color = "User type")
```


### Mobility diversity

```{r fig.height=8}
combinded_div <- div_inflow %>%
  st_set_geometry(NULL) %>%
  mutate(div_type = "Inflow") %>%
  bind_rows(.,
            div_outflow %>%
            st_set_geometry(NULL) %>%
            mutate(div_type = "Outflow")) %>%
  mutate(div_type = factor(div_type, levels = c("Outflow", "Inflow")))

combinded_div_mean <- combinded_div %>%
  group_by(div_type) %>%
  dplyr::summarise(mean = mean(norm_div_shannon)) %>%
  mutate(div_type = factor(div_type, levels = c("Outflow", "Inflow")))

ggplot(combinded_div, aes(norm_div_shannon, color = div_type, fill = div_type)) +
  geom_density(alpha = 0.5) +
  geom_vline(data = combinded_div_mean, aes(xintercept = mean, color = div_type), size = 1.5) +
  scale_x_continuous(breaks = seq(0, 1, 0.2)) +
  labs(x = "Normalized diversity", y = "Density", title = "Density distribution of normalized diversity", fill = "Diversity type", color = "Diversity type") +
  theme_bw() +
  theme(legend.position = c(0.1, 0.9))
```


## Regression analysis

### Inflow travel distance 

```{r}
## gather independent and dependent variables 
reg_inflow_dist <- mean_resale %>% 
  filter(grid_id %in% qualified_grids_inflow) %>% 
  left_join(., mean_dist_inflow %>% dplyr::select(-user_type)) %>% 
  left_join(., industrial_areas_grids %>% dplyr::select(grid_id, pct_industrial_area)) %>% 
  left_join(., residential_areas_grids %>% dplyr::select(grid_id, pct_residential_area)) %>%
  left_join(., dist2MRT_nearest %>% dplyr::select(grid_id, dist2MRT_km)) %>% 
  left_join(., df_pct_1_2room_rental %>% dplyr::select(grid_id, pct_1_2room_rental)) %>% 
  left_join(., dist2_center_grid %>% dplyr::select(grid_id, dist_grid2center_km)) %>% 
  replace(., is.na(.), 0)
head(reg_inflow_dist)
```


#### OLS

Formula: `Inflow distance ~ HDB resale price + Percentage of 1 ROOM and 2 Room rentals + Percentage of industrial building + Percentage of residential building + Distance to the nearest MRT station  + Distance to the central grid)`

```{r}
ols_inflow_dist <- lm(mean_dist_km ~ mean_resale_price_sqm + pct_1_2room_rental + pct_industrial_area + pct_residential_area + dist2MRT_km  + dist_grid2center_km, data = reg_inflow_dist)

summary(ols_inflow_dist)

spatial_viz(augment(ols_inflow_dist, data = reg_inflow_dist) %>% left_join(., grids) %>% st_as_sf(), 
            fill_var = ".resid", palette = "RdBu",
            legend_title = "Residual (OLS)", 
            main_title = "(a) Residuals of average incoming distance (Moran I statistic: 0.78)", 
            main.title_size = 1.6,
            legend.hist_height = 0.25, legend.hist_width = 0.5,
            legend_width = 0.4, legend.hist_size = 0.5,
            legend.title_size = 1.2, legend.text_size = 0.65)
```



#### Spatial Error Model (SEM)
```{r}
sf_reg_inflow_dist <- reg_inflow_dist %>% 
  left_join(., grids) %>% 
  st_as_sf() %>% 
  st_transform(crs = 4326)

dist_inflow_sp <- as(sf_reg_inflow_dist, 'Spatial')
dist_inflow_neighbors <- poly2nb(dist_inflow_sp)
summary(dist_inflow_neighbors)

dist_inflow_weights <- nb2listw(dist_inflow_neighbors, style="W", zero.policy=TRUE)
```


```{r}
moran.test(dist_inflow_sp$mean_dist_km, dist_inflow_weights)
```


```{r}
sem_inflow_dist <- spatialreg::errorsarlm(mean_dist_km ~ mean_resale_price_sqm + pct_1_2room_rental +  pct_industrial_area + pct_residential_area + pct_industrial_area + dist2MRT_km + dist_grid2center_km, data =  reg_inflow_dist, listw = dist_inflow_weights)
 
summary(sem_inflow_dist)
sf_reg_inflow_dist$resid_error <- residuals(sem_inflow_dist)

spatial_viz(sf_reg_inflow_dist,
            fill_var = "resid_error",
            legend_title = "Residual (SEM)", palette = "RdBu",
            main_title = "(a) Residuals of average incoming distance (Moran I statistic: - 0.09)",
            main.title_size = 1.6,
            legend.hist_height = 0.25, legend.hist_width = 0.5,
            legend_width = 0.4, legend.hist_size = 0.5,
            legend.title_size = 1.2, legend.text_size = 0.65)
moran.test(sf_reg_inflow_dist$resid_error, dist_inflow_weights)
```


#### Spatial Lag Model (SLM)
```{r}
slm_inflow_dist <- spatialreg::lagsarlm(mean_dist_km ~ mean_resale_price_sqm + pct_1_2room_rental + pct_industrial_area + pct_residential_area + dist2MRT_km + dist_grid2center_km, data =  reg_inflow_dist, listw = dist_inflow_weights)
summary(slm_inflow_dist)

sf_reg_inflow_dist$resid_lagsarlm <- residuals(slm_inflow_dist)
spatial_viz(sf_reg_inflow_dist,
            fill_var = "resid_lagsarlm",
            legend_title = "Residual (SLM)", 
            palette = "RdBu",
            main_title = "(a) Residuals of average incoming  distance (Moran I statistic: - 0.08)",
            main.title_size = 1.6,
            legend.hist_height = 0.25, legend.hist_width = 0.5,
            legend_width = 0.4, legend.hist_size = 0.5,
            legend.title_size = 1.2, legend.text_size = 0.65)
moran.test(sf_reg_inflow_dist$resid_lagsarlm, dist_inflow_weights)
```



```{r}
# compare three models
AIC(ols_inflow_dist, slm_inflow_dist, sem_inflow_dist)
```



### Outflow travel distance 

```{r}
reg_outflow_dist <- mean_resale %>% 
  filter(grid_id %in% qualified_grids_outflow) %>% 
  left_join(., mean_dist_inflow %>% dplyr::select(-user_type)) %>% 
  left_join(., industrial_areas_grids %>% dplyr::select(grid_id, pct_industrial_area)) %>% 
  left_join(., residential_areas_grids %>% dplyr::select(grid_id, pct_residential_area)) %>%
  left_join(., dist2MRT_nearest %>% dplyr::select(grid_id, dist2MRT_km)) %>% 
  # left_join(., pop_grids_2016) %>% 
  left_join(., df_pct_1_2room_rental %>% dplyr::select(grid_id, pct_1_2room_rental)) %>% 
  left_join(., dist2_center_grid %>% dplyr::select(grid_id, dist_grid2center_km)) %>% 
  replace(., is.na(.), 0) 
head(reg_outflow_dist)
```


#### OLS

Formula: `Outflow distance ~ HDB resale price + Percentage of 1 ROOM and 2 Room rentals + Percentage of industrial building + Percentage of residential building + Distance to the nearest MRT station + Distance to the central grid)`

```{r}
ols_outflow_dist <- lm(mean_dist_km ~ mean_resale_price_sqm + pct_1_2room_rental + pct_industrial_area + pct_residential_area + dist2MRT_km + dist_grid2center_km, data = reg_outflow_dist)

summary(ols_outflow_dist)

spatial_viz(augment(ols_outflow_dist, data = reg_outflow_dist) %>% left_join(., grids) %>% st_as_sf(), 
            fill_var = ".resid", palette = "RdBu",
            legend_title = "Residual (OLS)", 
            main_title = "(b) Residuals of average outgoing distance (Moran I statistic: 0.79) ",
            main.title_size = 1.6,
            legend.hist_height = 0.25, legend.hist_width = 0.5,
            legend_width = 0.4, legend.hist_size = 0.5,
            legend.title_size = 1.2, legend.text_size = 0.65)
```


#### Spatial Error Model (SEM)
```{r}
sf_reg_outflow_dist <- reg_outflow_dist %>% 
  left_join(., grids) %>% 
  st_as_sf() %>% 
  st_transform(crs = 4326)

dist_outflow_sp <- as(sf_reg_outflow_dist, 'Spatial')
dist_outflow_neighbors <- poly2nb(dist_outflow_sp)
summary(dist_outflow_neighbors)
dist_outflow_weights <- nb2listw(dist_outflow_neighbors, style="W", zero.policy=TRUE)
```


```{r}
moran.test(dist_outflow_sp$mean_dist_km, dist_outflow_weights)
```

```{r}
sem_outflow_dist <- spatialreg::errorsarlm(mean_dist_km ~ mean_resale_price_sqm + pct_1_2room_rental +  pct_industrial_area + pct_residential_area + dist2MRT_km + dist_grid2center_km, data =  reg_outflow_dist, listw = dist_outflow_weights)
 
summary(sem_outflow_dist)

sf_reg_outflow_dist$resid_error <- residuals(sem_outflow_dist)

spatial_viz(sf_reg_outflow_dist,
            fill_var = "resid_error",
            legend_title = "Residual (SEM)", palette = "RdBu",
            main_title = "(b) Residuals of average outgoing distance (Moran I statistic: -0.09)",
            main.title_size = 1.6,
            legend.hist_height = 0.25, legend.hist_width = 0.5,
            legend_width = 0.4, legend.hist_size = 0.5,
            legend.title_size = 1.2, legend.text_size = 0.65)
moran.test(sf_reg_outflow_dist$resid_error, dist_outflow_weights)
```

#### Spatial Lag Model (SLM)
```{r}
slm_outflow_dist <- spatialreg::lagsarlm(mean_dist_km ~ mean_resale_price_sqm + pct_1_2room_rental +  pct_industrial_area + pct_residential_area + dist2MRT_km +  dist_grid2center_km, data =  reg_outflow_dist, listw = dist_outflow_weights)
summary(slm_outflow_dist)

spatialreg::impacts(slm_outflow_dist, listw = dist_outflow_weights)

sf_reg_outflow_dist$resid_lagsarlm <- residuals(slm_outflow_dist)
spatial_viz(sf_reg_outflow_dist,
            fill_var = "resid_lagsarlm",
            legend_title = "Residual (SLM)", palette = "RdBu",
            main_title = "(b) Residuals of average outgoing distance (Moran I statistic: -0.08)",
            main.title_size = 1.6,
            legend.hist_height = 0.25, legend.hist_width = 0.5,
            legend_width = 0.4, legend.hist_size = 0.5,
            legend.title_size = 1.2, legend.text_size = 0.65)
moran.test(sf_reg_outflow_dist$resid_lagsarlm, dist_outflow_weights)
```

```{r}
# compare three models
AIC(ols_outflow_dist, slm_outflow_dist, sem_outflow_dist)
```



### Inflow diversity
```{r}
reg_inflow_div <- mean_resale %>% 
  filter(grid_id %in% qualified_grids_inflow) %>% 
  left_join(., div_inflow %>% st_set_geometry(NULL) %>% dplyr::select(grid_id, norm_div_shannon)) %>% 
  left_join(., industrial_areas_grids %>% dplyr::select(grid_id, pct_industrial_area)) %>% 
  left_join(., residential_areas_grids %>% dplyr::select(grid_id, pct_residential_area)) %>% 
  left_join(., dist2MRT_nearest %>% dplyr::select(grid_id, dist2MRT_km)) %>% 
  left_join(., df_pct_1_2room_rental %>% dplyr::select(grid_id, pct_1_2room_rental)) %>% 
  left_join(., dist2_center_grid %>% dplyr::select(grid_id, dist_grid2center_km)) %>%
  replace(., is.na(.), 0)

head(reg_inflow_div)
```

#### OLS 

```{r}
ols_inflow_div <- lm(norm_div_shannon ~ mean_resale_price_sqm + pct_1_2room_rental + pct_industrial_area + pct_residential_area + dist2MRT_km + dist_grid2center_km, data = reg_inflow_div)

summary(ols_inflow_div)

spatial_viz(augment(ols_inflow_div, data = reg_inflow_div) %>% left_join(., grids) %>% st_as_sf(), 
            fill_var = ".resid", palette = "RdBu",
            legend_title = "Residual (OLS)", 
            main_title = "(c) Residuals of inflow diversity (Moran I statistic: 0.69)",
            main.title_size = 1.6,
            legend.hist_height = 0.25, legend.hist_width = 0.5,
            legend_width = 0.4, legend.hist_size = 0.5,
            legend.title_size = 1.2, legend.text_size = 0.65)
```


#### Spatial Error (SEM) Models

```{r}
sf_reg_inflow_div <- reg_inflow_div %>% 
  left_join(., grids) %>% 
  st_as_sf() %>% 
  st_transform(crs = 4326)

div_inflow_sp <- as(sf_reg_inflow_div, 'Spatial')
div_inflow_neighbors <- poly2nb(div_inflow_sp)
summary(div_inflow_neighbors)
div_inflow_weights <- nb2listw(div_inflow_neighbors, style="W", zero.policy=TRUE)
```


```{r}
moran.test(div_inflow_sp$norm_div_shannon, div_inflow_weights)
```

```{r}
sem_inflow_div <- spatialreg::errorsarlm(norm_div_shannon ~ mean_resale_price_sqm + pct_1_2room_rental +  pct_industrial_area + pct_residential_area + dist2MRT_km + dist_grid2center_km, data =  reg_inflow_div, listw = div_inflow_weights)
summary(sem_inflow_div)

sf_reg_inflow_div$resid_error <- residuals(sem_inflow_div)
spatial_viz(sf_reg_inflow_div,
            fill_var = "resid_error",
            legend_title = "Residual (SEM)", palette = "RdBu",
            main_title = "(c) Residuals of inflow diversity (Moran I statistic: -0.05)",
            main.title_size = 1.6,
            legend.hist_height = 0.25, legend.hist_width = 0.5,
            legend_width = 0.4, legend.hist_size = 0.5,
            legend.title_size = 1.2, legend.text_size = 0.65)
moran.test(sf_reg_inflow_div$resid_error, div_inflow_weights)
```

#### Spatial Lag Models (SLM):
```{r}
slm_inflow_div <- spatialreg::lagsarlm(norm_div_shannon ~ mean_resale_price_sqm + pct_1_2room_rental + pct_industrial_area + pct_residential_area + dist2MRT_km +dist_grid2center_km, data =  reg_inflow_div, listw = div_inflow_weights)
summary(slm_inflow_div)
spatialreg::impacts(slm_inflow_div, listw = div_inflow_weights)

sf_reg_inflow_div$resid_lagsarlm <- residuals(slm_inflow_div)

spatial_viz(sf_reg_inflow_div,
            fill_var = "resid_lagsarlm",
            legend_title = "Residual (SLM)", palette = "RdBu",
            main_title = "(c) Residuals of inflow diversity (Moran I statistic: -0.03)",
            main.title_size = 1.6,
            legend.hist_height = 0.25, legend.hist_width = 0.5,
            legend_width = 0.4, legend.hist_size = 0.5,
            legend.title_size = 1.2, legend.text_size = 0.65)
moran.test(sf_reg_inflow_div$resid_lagsarlm, div_inflow_weights)
```


```{r}
AIC(ols_inflow_div, slm_inflow_div, sem_inflow_div)
```



### Outflow diversity 

```{r}
reg_outflow_div <- mean_resale %>% 
  filter(grid_id %in% qualified_grids_outflow) %>% 
  left_join(., div_outflow %>% st_set_geometry(NULL) %>% dplyr::select(grid_id, norm_div_shannon)) %>% 
  left_join(., industrial_areas_grids %>% dplyr::select(grid_id, pct_industrial_area)) %>% 
  left_join(., residential_areas_grids %>% dplyr::select(grid_id, pct_residential_area)) %>% 
  left_join(., dist2MRT_nearest %>% dplyr::select(grid_id, dist2MRT_km)) %>% 
  left_join(., df_pct_1_2room_rental %>% dplyr::select(grid_id, pct_1_2room_rental)) %>% 
  left_join(., dist2_center_grid %>% dplyr::select(grid_id, dist_grid2center_km)) %>%
  replace(., is.na(.), 0) 
head(reg_inflow_div)
```

#### OLS 

```{r}
ols_outflow_div <- lm(norm_div_shannon ~ mean_resale_price_sqm + pct_1_2room_rental + pct_industrial_area + pct_residential_area + dist2MRT_km + dist_grid2center_km, data = reg_outflow_div)
summary(ols_outflow_div)

spatial_viz(augment(ols_outflow_div, data = reg_outflow_div) %>% left_join(., grids) %>% st_as_sf(), 
            fill_var = ".resid", palette = "RdBu",
            legend_title = "Residual (OLS)", 
            main_title = "(d) Residuals of outflow diversity (Moran I statistic: 0.83)",
            main.title_size = 1.6,
            legend.hist_height = 0.25, legend.hist_width = 0.5,
            legend_width = 0.4, legend.hist_size = 0.5,
            legend.title_size = 1.2, legend.text_size = 0.65)
```


```{r}
sf_reg_outflow_div <- reg_outflow_div %>% 
  left_join(., grids) %>% 
  st_as_sf() %>% 
  st_transform(crs = 4326)

div_outflow_sp <- as(sf_reg_outflow_div, 'Spatial')
div_outflow_neighbors <- poly2nb(div_outflow_sp)
summary(div_outflow_neighbors)
div_outflow_weights <- nb2listw(div_outflow_neighbors, style="W", zero.policy=TRUE)
```


```{r}
moran.test(div_outflow_sp$norm_div_shannon, div_outflow_weights)
```


#### Spatial Error (SEM) Models
```{r}
sem_outflow_div <- spatialreg::errorsarlm(norm_div_shannon ~ mean_resale_price_sqm + pct_1_2room_rental +  pct_industrial_area + pct_residential_area + dist2MRT_km + dist_grid2center_km, data =  reg_outflow_div, listw = div_outflow_weights)
summary(sem_outflow_div)

sf_reg_outflow_div$resid_error <- residuals(sem_outflow_div)
spatial_viz(sf_reg_outflow_div,
            fill_var = "resid_error",
            legend_title = "Residual (SEM)", palette = "RdBu",
            main_title = "(d) Residuals of outflow diversity (Moran I statistic: -0.08)",
            main.title_size = 1.6,
            legend.hist_height = 0.25, legend.hist_width = 0.5,
            legend_width = 0.4, legend.hist_size = 0.5,
            legend.title_size = 1.2, legend.text_size = 0.65)
moran.test(sf_reg_outflow_div$resid_error, div_outflow_weights)
```


#### Spatial Lag Models (SLM):
```{r}
slm_outflow_div <- spatialreg::lagsarlm(norm_div_shannon ~ mean_resale_price_sqm + pct_1_2room_rental +  pct_industrial_area + pct_residential_area + dist2MRT_km + dist_grid2center_km, data =  reg_outflow_div, listw = div_outflow_weights)
summary(slm_outflow_div)
spatialreg::impacts(slm_outflow_div, listw = div_outflow_weights)

sf_reg_outflow_div$resid_lagsarlm <- residuals(slm_outflow_div)
spatial_viz(sf_reg_outflow_div,
            fill_var = "resid_lagsarlm",
            legend_title = "Residual (SLM)", palette = "-RdBu",
            main_title = "(d) Residuals of outflow diversity (Moran I statistic: -0.06)",
            main.title_size = 1.6,
            legend.hist_height = 0.25, legend.hist_width = 0.5,
            legend_width = 0.4, legend.hist_size = 0.5,
            legend.title_size = 1.2, legend.text_size = 0.65)
moran.test(sf_reg_outflow_div$resid_lagsarlm, div_outflow_weights)
```

```{r}
AIC(ols_outflow_div, slm_outflow_div, sem_outflow_div)
```
